---
title: "lab 6"
author: "Ruoyu Wang"
date: "2020/2/13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
# attach packages
library(tidyverse)
library(here)
library(janitor)
library(raster)
library(sf)
library(tmap)
library(tmaptools)
library(gstat)
```

## grand canyon DeoTIFF

```{r}
gc_dem <- raster::raster(here::here("data", "gc_dem.tif"))

# look at it using plot()
# plot(gc_dem)

# Check the CRS: coordinate system
# gc_dem@crs
# it's wired because it shows in meters not long/latitude

# check the extent (bounding box):
# gc_dem@extent

# update crs, creating a wgs84 w/latlong:
wgs84 = "+proj=longlat +datum=WGS84 +ellps=WGS84 +no_defs" # Just have this ready to copy/paste

gc_reproj = projectRaster(gc_dem, crs = wgs84, method = "bilinear")
  
gc_reproj@extent
```

### crop raster to a smaller area (bounding box):
```{r}
bounds <- as(extent(-122.4, -112.0, 36.1, 36.3), 'SpatialPolygons')
```

### Make the CRS of our bounding box the same as thr gc_reproj
```{r}
crs(bounds) <- crs(gc_reproj)
```

### now le's crop our original sparial data:
```{r}
gc_crop <- crop(gc_reproj, bounds)

plot(gc_crop)
```

### Resample using the `aggregate` function:
```{r}
gc_agg <- raster::aggregate(gc_crop, fact = 50)
gc_agg2 <- raster::aggregate(gc_crop, fact = 10)

plot(gc_agg)
plot(gc_agg2)
```

### now let's get safe and sound in ggplot
first, convert data to a data frame:
```{r}
gc_df <- as.data.frame(gc_agg, xy=TRUE)

ggplot(data = gc_df, aes(x = x, y = y))+
  geom_raster(aes(fill = gc_dem))+
  coord_quickmap()+
  theme_minimal()+
  scale_fill_gradientn(colors = c(
    "purple", "red","magenta","yellow","white"
  ))
```

### How can I just select cells that match viven criteria?

i have some species of vegatation that will only grow when gc_dem is between 1000 and 1500

```{r}
gc_hab <- gc_crop

# set any cells outside of (1000 - 1500) to NA
gc_hab[gc_hab > 1500 | gc_hab < 1000] <- NA

plot(gc_hab)

```

now let's make this interactive with tmap:
```{r}
tmap_mode("view")

tm_shape(gc_hab)+
  tm_raster(legend.show = FALSE, palette = "plasma")
```

### kriging rain in kansas

Read in the KS counties shapefile data
```{r}
ks_counties <- read_sf(here::here("data", "ks_counties", "ks_counties_shapefile.shp"))

# use base plot:
plot(ks_counties)

# Check CRS
st_crs(ks_counties) # no crs, we should set one

# set to EPSG 4326:
st_crs(ks_counties) <- 4326

plot(ks_counties)
```

practice in ggplot:
```{r}
ggplot(data = ks_counties)+
  geom_sf()
```


Now let's read in the rainfall data
```{r}
ks_rain <- read_csv(here("data", "ks_rain.csv")) %>% 
  clean_names()
```

update the ks_rain data to be recognized as spatial points:
```{r}
ks_sf <- st_as_sf(ks_rain, coords = c("lon", "lat"), crs = 4326) # has to be in long/lat order
```


```{r}
ggplot() +
  geom_sf(data = ks_counties)+
  geom_sf(data = ks_sf,
          aes(color = amt, size = amt),
          show.legend = FALSE)+
  theme_minimal()
```

## kriging to predict rainfall:
```{r}
ks_sp <- as_Spatial(ks_sf)
class(ks_sp)
```

### make a spatial pixels grid that we'll make predictions over:
```{r}
# bbox(ks_sp)
#   min    max
# coords.x1 -101.75 -94.63
# coords.x2   37.00  40.00

lat <- seq(37, 40, length.out = 200)
long <- seq(-94.6, -102, length.out = 200)

# now we're going to make this into a spatial grid:
grid <- expand.grid(lon = long, lat = lat)
grid_sf <- st_as_sf(grid, coords = c("lon", "lat"), crs = 4326)
grid_sp <- as_Spatial(grid_sf)

plot(grid_sp)
```

### Now make a variogram
```{r}
ks_vgm <- variogram(amt ~ 1, data = ks_sp)
plot(ks_vgm)

# My estimates for variogram parameters:
# nugget = 0.2
# sill = 0.8
# range = 200

ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.1, psill = 0.8, range = 200, model = "Sph")) # other models: "Exp", "Gau"

ks_vgm_fit
# nug: 0.102, sill: 0.954, range: 235.142

plot(ks_vgm, ks_vgm_fit)
```

### now, krige!
```{r}
ks_krige <- krige(amt ~ 1, ks_sp, grid_sp, model = ks_vgm_fit)

ks_krige@data
# for each cell, it has a predicted value and variances
```

```{r}
spplot(ks_krige, "var1.pred")
# cute with a heart!
```

### Make a data frame of kriged predictions
```{r}
ks_df <- data.frame(ks_krige@data["var1.pred"],
                    ks_krige@data["var1.var"],
                    ks_krige@coords) %>% 
  rename(longitude = coords.x1,
         latitude = coords.x2)

# converting this data object
rain_sf <- st_as_sf(ks_df, coords = c("longitude", "latitude"), crs = 4326)

ggplot(rain_sf)+
  geom_sf(aes(color = var1.pred))
```

# crop to the actual KS outline
```{r}
ks <- read_sf(dsn = here::here("data", "states"), layer = "cb_2017_us_state_20m") %>% 
  dplyr::select(NAME) %>% 
  filter(NAME == "Kansas") %>% 
  st_transform(crs = 4326)

plot(ks)
```

```{r}
# find the intersection of the two:
rain_sf_ks <- st_intersection(rain_sf, ks)

# plot
ggplot(data = rain_sf_ks)+
  geom_sf(aes(color = var1.pred))
```




